Forecasting the US Election 2024

### 2024-11-03 paper/paper.qmd
Updated paper with toc, abstract, draft results and discussion.

---
title: "Forecasting the US Election 2024"
subtitle: "##TODO: result here"
author: 
  - Maryam Ansari
  - Amy Jin
  - Maggie Zhang

thanks: "Code and data are available at: [https://github.com/aj3616/Forcasting-US-Elections](https://github.com/aj3616/Forcasting-US-Elections)."
date: today
date-format: long
abstract: "first sentence: specify the general area of the paper and encourage the reader; second sentence: specify the dataset and methods at a general level; third sentence: specify the headline result; and a fourth sentence about implications"
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Our results are summarized in @tbl-modelresults.

Results will likely require summary statistics, tables, graphs, images, and possibly statistical analysis or maps. There should also be text associated with all these aspects.

Show the reader the results by plotting them where possible. Talk about them. Explain them. That said, this section should strictly relay results.

Regression tables must not contain stars.

@@

## What is done in this paper? {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## What is something that we learn about the world?

Please don't use these as sub-heading labels - change them to be what your point actually is.

## What is another thing that we learn about the world?

## Weaknesses and next steps

What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?

Weaknesses and next steps should also be included.

>>>>

---
title: "Forecasting the US Election 2024"
subtitle: "Modeling the U.S. Polling Data to forecast the outcome of he 2024 U.S. Presidential Elections"
author: 
  - Maryam Ansari
  - Amy Jin
  - Maggie Zhang

thanks: "Code and data are available at: [https://github.com/aj3616/Forcasting-US-Elections](https://github.com/aj3616/Forcasting-US-Elections)."
date: today
date-format: long
abstract: "This paper uses polling data to project the outcome of the 2024 U.S. Presidential election between Kamala Harris and Donald Trump. By employing a multiple linear regression model, we predict each candidate’s level of support across states and estimate the corresponding electoral vote counts. Our analysis suggests that Harris has a potential advantage in the Electoral College, highlighting the influence of polling trends and pollster-specific biases on election forecasting. These findings underscore the importance of aggregating diverse polling data to create a more balanced and accurate prediction of election results."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

To project the outcome of the 2024 U.S. Presidential election, we utilized a statistical model that leverages polling data from FiveThirtyEight to predict overall support for each candidate. Our approach employs a regression model that incorporates poll end dates and pollster-specific adjustments, enabling us to generate a "poll-of-polls" forecast and relay these results as a predictor of election outcomes.

**1. Prediction of Polling-Based Support:** Using Model 2, we estimated support levels by regressing polling outcomes on poll end dates and pollster-specific indicators. This model was applied to generate predicted support percentages for each candidate over the polling period, accounting for variations across pollsters (see Table 1 for regression output). These estimates reflect the predicted level of support for each candidate as the election approaches, adjusted for biases that are characteristic of individual polling organizations.

**2. Time-Based Trend Analysis:** By incorporating poll end dates, our model captures temporal trends, allowing us to observe shifts in candidate support over time. Figure 1 displays the predicted support levels for each candidate over the final months leading up to the election, highlighting any momentum or shifts that may favor one candidate. This visualization provides insight into how public opinion has evolved closer to the election date.

**3. Adjustment for Pollster-Specific Biases:** Pollster biases can significantly influence polling averages. Our model addresses these variations through the pollster-specific coefficients \( \gamma_j \), which adjust each pollster's results based on historical tendencies. Figure 2 illustrates these adjustments, showing how certain pollsters tend to lean in favor of one candidate over another. This correction improves the reliability of our aggregated forecast by balancing these systematic biases.

**4. Estimation of Overall Support:** Aggregating the predicted support from all polls, we computed an average prediction for each candidate as election day nears. This aggregation provides a summary of anticipated national-level support based on polling data and adjusted for timing and pollster effects (see Table 2 for summary statistics of aggregated support levels). This result allows for a more precise comparison of each candidate's standing in the final stretch of the campaign.

**5. Interpretation of Results:** The analysis suggests that [Candidate A] holds a lead in projected support, with a predicted margin of [X percentage points] over [Candidate B]. This result implies a favorable position for [Candidate A], assuming that polling trends continue as projected. While this forecast is based on aggregate polling data, the adjustments for time trends and pollster effects lend robustness to the model's predictions, offering a refined projection of likely support levels for each candidate.

This methodology effectively combines poll end dates and pollster adjustments, bridging raw polling data with refined predictions of national support. By adjusting for bias and temporal effects, our model provides a comprehensive forecast that can inform expectations about the election outcome.

@@

## Summary of Findings
This study utilizes polling data to project the results of the 2024 U.S. Presidential election between Kamala Harris and Donald Trump. We developed a multiple linear regression model that predicts state-level polling outcomes by incorporating the end date of each poll and accounting for differences in polling methodology across various organizations. By aggregating state-level predictions and mapping them to the U.S. Electoral College framework, we were able to estimate the potential electoral vote count for each candidate. Our model suggests that Kamala Harris has an advantage in terms of projected electoral votes, signaling a possible pathway to victory based on the current polling trends. These findings underscore the effectiveness of using a ‘poll-of-polls’ approach and highlight the importance of controlling for both time trends and pollster-specific biases in forecasting.

## Insights from results
Several key insights emerge from our analysis. First, we observe that poll end dates significantly influence the model’s predictions, indicating that voter preferences are not static but fluctuate over time. This temporal sensitivity suggests that the timing of polling data collection is crucial for accuracy, particularly as Election Day approaches. Second, the model's adjustment for pollster-specific effects reveals substantial variation among polling organizations, with some systematically showing higher support for one candidate over the other. This finding highlights the necessity of accounting for methodological differences between pollsters to avoid skewed predictions. By integrating these pollster-specific adjustments, we enhance the model’s robustness and create a more balanced and comprehensive projection of voter support.

## Implications of the results
The implications of these findings are twofold. On one level, our results underscore the volatility of public opinion and the influence of polling timing on the perceived support for each candidate. The sensitivity to poll timing means that election forecasts can shift rapidly, particularly in the final weeks before the election. This insight could be valuable for political campaigns and analysts, who must consider the timing of major announcements or policy shifts to capture late-breaking trends in voter sentiment. On another level, the variation among pollsters suggests that analysts should be cautious when interpreting individual poll results. Relying on a single poll or even a small set of polls could lead to biased conclusions; aggregating across multiple polls and accounting for systematic differences is essential for obtaining a balanced view.

## Future Communication and Work to Be Done
While our model provides a strong foundation for predicting electoral outcomes, there are areas where future research could expand and refine these methods. One potential improvement would be to incorporate additional variables, such as economic indicators (e.g., unemployment rates, inflation) or significant political events (e.g., debates, foreign policy crises) that may affect voter sentiment. Including real-time data feeds could also allow for dynamic model updates, enabling continuous refinement of predictions as new polls are released and conditions evolve.

Another promising direction for future work involves exploring machine learning techniques, such as ensemble models, that could capture complex interactions between variables and improve predictive power. Additionally, understanding the impact of digital influence—particularly social media—on voter behavior might provide further context for shifts in public opinion. Expanding the model in this way could offer deeper insights into the drivers of voter sentiment and enhance the accuracy of future election forecasts.

Finally, there is room for improving the communication of model results to the public and stakeholders. Visualizations of the model’s uncertainty, for example, could help clarify the inherent unpredictability of election forecasting and avoid overconfident interpretations of the results. Regular updates or a publicly accessible dashboard could also foster greater transparency, providing a real-time window into how predictions shift with new data. This open communication would not only make election predictions more accessible to the general public but also underscore the value of aggregating diverse data sources in political forecasting.
